Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 casestudies.tex
some work

In this section, we present a simple, but real-world example
application currently deployed using RCloud.

This example application is an implementation of LDAVis, a
recently-published method for visualizing large collections of
text. It was specifically designed to enable non-expert users to
explore collections of short text documents using topic modeling and
visualization. Topic modeling~\cite{Blei:2003:LDA} is a standard
technique for text summarization which, although powerful, requires a
certain amount of manual intervention and
interpretation~\cite{Sievert:2014:LAM}.

@@

LDAVis itself was developed by two members of the technical staff at
AT\&T Labs, and originally targeted RStudio's
Shiny~\cite{RStudio:2013:SWA}, a framework for developing R
applications for the web.  While Shiny provides unparalleled ease of
development, discoverability and deployment turn out to be equally
important aspects in the lifecycle of an internal application (see
Section~\ref{sec:interviews}).  On these aspects, RCloud provides a
very simple model: \emph{all developed notebooks are automatically
  deployed}.

LDAVis combines textual analysis, dimensionality reduction, and
interactive visualization. The text analysis is performed via a
combination of an existing R library that fits an LDA model using
Gibbs sampling~\cite{} and some R code written specifically by the
developers. The analysis module is, ultimately, a single R function
that is exposed to the web application via the Javascript-R RPC
mechanism described in Section~\ref{sec:system}; this way, the
analysis is executed remotely on the RCloud servers.

Each textual topic is a probability distribution over the all the
words of the document. In order to expose patterns in the
relationships between the topics, LDAVis employs a combination of
interactive visualization and dimensionality reduction (letting users
choose different measures for the topic distances and different
dimensionality reduction techniques). The dimensionality reduction
algorithms and distance measures are again implemented in R, which
means they are executed on the RCloud servers as well.

The result of the dimensionality reduction is a two-dimensional plot
of the topic space, which can be seen in
Figure~\ref{fig:textvis}. Describe d3 view.

This application is simple, but highlights some of the unique features
of RCloud. While RCloud notebooks allow deployment of R analyses over
the web with no additional effort, RCloud \emph{applications} are more
powerful, and are developed in a combination of Javascript and HTML
for the front end. This requires additional knowledge over Shiny, but
we argue that the RCloud model makes the analysis side simpler for the
analysts (since they simply write the R code in the style they are
used to), and the front-end visualization side simpler for the
front-end developers (since they simply write the Javascript code in
the style \emph{they} are used to). In addition, RCloud applications
inherit the automatic deployment and discoverability features of
regular RCloud notebooks.

>>>>

As an example, we present a simple real-world application
deployed under RCloud.

The application is an implementation of LDAVis,
a recent method for visualizing large text corpora.
It was designed to help non-experts to explore
collections of short text documents using topic modeling and
visualization. Topic modeling~\cite{Blei:2003:LDA} is a standard
technique for text summarization which, although powerful, requires
some manual intervention and interpretation~\cite{Sievert:2014:LAM}.

@@

LDAVis was developed by two technical staff members at
AT\&T Labs, and originally was written in RStudio's
Shiny~\cite{RStudio:2013:SWA}, a framework for developing R
applications for the web.  While Shiny provides outstanding ease of
development, it turns out that discoverability and deployment are equally
important aspects in the lifecycle of an internal application (see
Section~\ref{sec:interviews}). In these aspects, RCloud provides a
simple solution: \emph{all developed notebooks are automatically deployed}.

LDAVis combines text analysis, dimensionality reduction and
interactive visualization. Text analysis is performed by
combining a standard R library to fit LDA models using
Gibbs sampling~\cite{} with custom R code written by the
developers. The analysis module is, ultimately, a single R function
exposed to the web application via the Javascript-R RPC
mechanism described in Section~\ref{sec:system};
thus analysis is performed remotely on an RCloud server.

Each text topic is a probability distribution over the all the
words of the document. To expose patterns in the relationships
between topics, LDAVis employs a combination of
interactive visualization and dimensionality reduction,
allowing users to adjust measures for topic distances and the choice
of dimensionality reduction technique. The dimensionality reduction
algorithms and distance measures are implemented in R, which
means they can be executed on the RCloud servers as well.

The result of the dimensionality reduction process is a
two-dimensional plot of the topic space, an example of which is shown in
Figure~\ref{fig:textvis}. The interactive view is implemented in SVG
and Javascript through D3~\cite{Bostock:2011:DDD}. One of the most
popular web-based visualization libraries for R is ggvis~\cite{ggvis},
so it is natural to ask if ggvis could have been used instead of
custom Javascript. In this case, the interactive features of ggvis
(and, by extension, Shiny) are a subset of those in Vega~\cite{vega}.
Custom interactions in LDAVis (such as hovering over a topic,
topic cluster, or word) do not appear to be available yet in Vega~\cite{vega},
although the required components for reactive interaction were
recently described by Satyanarayan et
al.~\cite{Satyanarayan:2014:DID}.
Although custom Javascript was required, the flexibility is welcomed
by many web developers.

The LDAVis application is simple, but highlights some unique features
of RCloud. While RCloud notebooks allow deployment of R analyses over
the web with no additional effort, RCloud \emph{applications} are more
powerful, and are developed with a combination of Javascript and HTML
for the front end. This requires additional knowledge over Shiny, but
we argue that the RCloud model makes the analysis side simpler 
(since analysts simply write R in the style they are used to),
and the front end visualization side is simpler for the web
developers (since they simply write the Javascript code in
the style \emph{they} are used to). In addition, RCloud applications
inherit the automatic deployment and discoverability features of
all RCloud notebooks.
